% Platform Economics: Pre-Class Intro (3 slides)
% Shows main output and learning achievements after the lecture
\input{../template_beamer.tex}

% ======================= DOCUMENT INFO =======================
\title[Lecture Preview]{Platform Economics --- What You Will Learn}
\subtitle{A Preview of the Lecture Outcomes}
\author{Joerg Osterrieder}
\institute{Digital Finance}
\date{}

\begin{document}

% =======================================================================
% SLIDE 1: The Big Picture
% =======================================================================
\begin{frame}{Platform Economics: The Big Picture}

\begin{alertblock}{After this lecture, you will see digital markets differently.}
\end{alertblock}

\vspace{2mm}
\footnotesize
Platforms are the dominant business model of the digital economy. They power payments, lending, trading, and insurance --- yet most people cannot explain \emph{why} they work or \emph{when} they fail.

\vspace{3mm}
\textbf{\color{mlpurple}This lecture gives you a complete analytical toolkit:}

\vspace{2mm}
\begin{columns}[T]
\begin{column}{0.48\textwidth}
\begin{block}{Theory}
\begin{itemize}\compactlist
\item What makes a platform different from a traditional business
\item How network effects create explosive growth --- and collapse
\item Why some markets tip to one winner while others stay competitive
\end{itemize}
\end{block}
\end{column}
\begin{column}{0.48\textwidth}
\begin{block}{Practice}
\begin{itemize}\compactlist
\item How platforms price, launch, and compete
\item How to evaluate whether a FinTech has a real moat
\item How regulation and decentralization are reshaping the landscape
\end{itemize}
\end{block}
\end{column}
\end{columns}

\vspace{3mm}
\begin{exampleblock}{Format}
\scriptsize 8 sections | 59 frames | 13 diagrams | 4 case discussions | 2 self-assessments | 1 group workshop
\end{exampleblock}
\end{frame}

% =======================================================================
% SLIDE 2: Six Things You Will Be Able To Do
% =======================================================================
\begin{frame}{After the Lecture: Six Things You Will Be Able To Do}
\footnotesize

\begin{enumerate}\compactlist
\item \textbf{Distinguish platforms from pipelines} --- instantly classify any business by asking: ``Does it \emph{connect} or does it \emph{produce}?''

\vspace{1mm}
\item \textbf{Identify all four types of network effects} --- direct, indirect, data, and negative --- and explain how each shapes competition

\vspace{1mm}
\item \textbf{Diagnose the chicken-and-egg problem} and pick from six proven launch strategies (subsidize, single-player mode, seed supply, piggyback, marquee user, micro-market)

\vspace{1mm}
\item \textbf{Evaluate unit economics} using CAC, LTV, payback period, and churn --- and spot when growth is real vs.\ venture-subsidized

\vspace{1mm}
\item \textbf{Apply a six-question framework} to any FinTech platform: network effects? unit economics? switching costs? data advantage? defensibility? regulation?

\vspace{1mm}
\item \textbf{Debate the future}: centralized vs.\ decentralized platforms, token economics, AI-driven network effects, and the regulator's dilemma
\end{enumerate}

\vspace{2mm}
\begin{alertblock}{The Core Skill}
\scriptsize Analyze any platform business and assess whether it has a sustainable competitive advantage --- or just expensive user acquisition.
\end{alertblock}
\end{frame}

% =======================================================================
% SLIDE 3: Key Frameworks You Will Take Away
% =======================================================================
\begin{frame}{Key Frameworks You Will Take Away}
\footnotesize

\begin{columns}[T]
\begin{column}{0.48\textwidth}

\textbf{\color{mlpurple}1. Platform Classification}\\[1mm]
\scriptsize Pipeline vs.\ platform; transaction, innovation, investment, integrated types.

\vspace{3mm}
\textbf{\color{mlpurple}2. Network Effects Toolkit}\\[1mm]
\scriptsize Four types, critical mass S-curve, winner-take-most conditions (network effects + switching costs + low multi-homing).

\vspace{3mm}
\textbf{\color{mlpurple}3. Pricing Logic}\\[1mm]
\scriptsize Subsidy side vs.\ money side; Rochet-Tirole pricing principle; access, usage, and enhanced fees.

\end{column}
\begin{column}{0.48\textwidth}

\textbf{\color{mlpurple}4. Business Model Assessment}\\[1mm]
\scriptsize Commission, subscription, freemium, advertising, PFOF, float; unit economics health check.

\vspace{3mm}
\textbf{\color{mlpurple}5. Data Flywheel}\\[1mm]
\scriptsize More users $\rightarrow$ more data $\rightarrow$ better algorithms $\rightarrow$ better experience $\rightarrow$ more users. Why this moat compounds.

\vspace{3mm}
\textbf{\color{mlpurple}6. Sustainability Test}\\[1mm]
\scriptsize Six-question evaluation framework applicable to any FinTech you encounter --- in this course or in your career.

\end{column}
\end{columns}

\vspace{4mm}
\begin{exampleblock}{Ready?}
\scriptsize These frameworks will let you move from ``I use platforms'' to ``I understand how platforms work, compete, and fail.''
\end{exampleblock}

\bottomnote{Lecture based on Rochet-Tirole (2003/2006), Parker-Van Alstyne-Choudary (2016), Eisenmann-Parker-Van Alstyne (2006), Evans-Schmalensee (2016).}
\end{frame}

\end{document}
