% Platform Economics: Pre-Class Discovery Handout
% 2-page handout to prime student thinking before the lecture
\input{../template_beamer.tex}

% ======================= DOCUMENT INFO =======================
\title[Pre-Class Discovery]{Platform Economics --- Pre-Class Discovery}
\subtitle{Think About These Before We Meet}
\author{Joerg Osterrieder}
\institute{Digital Finance}
\date{}

\begin{document}

% =======================================================================
% PAGE 1: Everyday Platforms -- What Do You Already Know?
% =======================================================================
\begin{frame}{Everyday Platforms: What Do You Already Know?}

\begin{alertblock}{You already use platforms every day --- but have you ever thought about \emph{why} they work?}
\end{alertblock}

\vspace{2mm}
\footnotesize
\textbf{\color{mlpurple}Discovery Activity 1: Your Platform Diary}

Think about the last 48 hours. List every digital service you used where \textbf{two different groups} interact through the same app or website (e.g., buyers and sellers, riders and drivers, borrowers and lenders).

\vspace{2mm}
\begin{columns}[T]
\begin{column}{0.48\textwidth}
\begin{block}{For each service, note:}
\begin{itemize}\compactlist
\item Who are the two (or more) sides?
\item Does the service \emph{make} something, or does it \emph{connect} people?
\item Could one side exist without the other?
\item Did you pay? If not, who does?
\end{itemize}
\end{block}
\end{column}
\begin{column}{0.48\textwidth}
\begin{block}{Then ask yourself:}
\begin{itemize}\compactlist
\item Why do you use \emph{this} service and not an alternative?
\item Would it be easy to switch? Why or why not?
\item Is the service more useful because many people use it?
\item What would happen if half the users left overnight?
\end{itemize}
\end{block}
\end{column}
\end{columns}

\vspace{3mm}
\begin{exampleblock}{Bring to class}
\scriptsize Your list of 3--5 services and your answers. We will use these as examples throughout the lecture.
\end{exampleblock}

\bottomnote{No right or wrong answers --- the goal is to notice patterns you already experience but may not have named yet.}
\end{frame}

% =======================================================================
% PAGE 2: Three Puzzles to Ponder
% =======================================================================
\begin{frame}{Three Puzzles to Ponder}
\footnotesize

\textbf{\color{mlpurple}Puzzle 1: The Empty Marketplace}
\begin{quote}
\scriptsize
You want to launch a new online marketplace connecting local farmers with restaurant chefs. On day one, you have zero farmers and zero chefs. Farmers won't list produce if no chefs are browsing. Chefs won't browse if no produce is listed. \textbf{How do you escape this trap?} Write down at least two ideas.
\end{quote}

\vspace{2mm}
\textbf{\color{mlpurple}Puzzle 2: The Free Product}
\begin{quote}
\scriptsize
A popular financial app lets you buy and sell stocks with zero commission. The company has thousands of employees, expensive offices, and sophisticated technology. \textbf{If you are not paying, how does the company make money?} Who \emph{is} paying, and why? Is ``free'' really free?
\end{quote}

\vspace{2mm}
\textbf{\color{mlpurple}Puzzle 3: The Unbeatable Incumbent}
\begin{quote}
\scriptsize
A dominant payment network is accepted at virtually every store. A new competitor offers lower fees and better technology. Yet almost no merchants switch. \textbf{Why not?} What invisible force keeps merchants locked in, even when a cheaper option exists?
\end{quote}

\vspace{2mm}
\begin{alertblock}{Come prepared}
\scriptsize Jot down your best guesses. In the lecture, we will name the economic forces behind each puzzle and give you frameworks to analyze any platform business.
\end{alertblock}
\end{frame}

\end{document}
